---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 1000
)
```

# climateapi

<!-- badges: start -->
<!-- badges: end -->

The goal of `library(climateapi)` is to minimize repeated data cleaning and wrangling
to enable project teams to devote more time to substantive analysis and inference-making.
The package works toward this goal by creating a unified interface to common datasets
and data manipulation tasks. Functions (will) support both climate-specific datasets
as well as those that are climate-adjacent.

## Installation

You can install the development version of climateapi from [GitHub](https://github.com/) with:

``` r
# install.packages("renv")
renv::install("UI-Research/climateapi")
```

## The `climateapi` package at work:

```{r setup}
library(climateapi)
#renv::install("UI-Research/urbnindicators")
#renv::install("UrbanInstitute/urbnthemes")
library(urbnindicators)
library(sf)
library(tidyverse)
library(urbnthemes)
set_urbn_defaults(style = "print")
```

```{r example}
county_disaster_declarations = get_fema_disaster_declarations_county(api = TRUE)

county_disaster_declarations %>%
  filter(stringr::str_detect(GEOID, "^01")) %>% ## Alabama
  group_by(year_declared) %>%
    summarize(annual_incidents = sum(incidents_all, na.rm = TRUE)) %>%
  ggplot() +
    geom_col(aes(x = year_declared, y = annual_incidents)) +
    labs(
      title = "COVID Results in a Spike of Counties with Disaster Declarations in 2020",
      subtitle = "Sum of major disaster declarations per Alabama county, by year",
      x = "",
      y = "") +
    theme_urbn_print()
```

```{r}
wildfires_la = get_current_fire_perimeters(
  bbox = tigris::counties(cb = TRUE, state = "CA") %>%
    dplyr::filter(stringr::str_detect(NAMELSAD, "Angeles")) %>%
    sf::st_bbox())

la_impacted_structues = estimate_impacted_structures(
  boundaries = wildfires_la,
  geography = "tract")

tracts_sf = tigris::tracts(cb = TRUE, year = 2023) %>%
  sf::st_filter(wildfires_la %>% st_buffer(10000)) %>%
  dplyr::select(GEOID) %>%
  dplyr::left_join(la_impacted_structures, by = "GEOID")

ggplot() +
  geom_sf(data = tracts_sf)
```

