---
title: "Preliminary Damage Assessments and Declaration Outcomes"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    fig-width: 8
    fig-height: 6
    fig-dpi: 300
execute:
  echo: true
  ## warnings and messages are shown here, unlike in the package vignette:
  ## `get_preliminary_damage_assessments()` reports its data quality problems and
  ## its unmatched reports that way, and when you are stepping through the
  ## analysis those are things you want to read rather than suppress
  warning: true
  message: true
editor_options:
  chunk_output_type: console
---

```{r}
#| label: options
#| include: false

options(scipen = 999)
```

## Overview

When a state, territory, or tribal government believes an event exceeds its
capacity to respond, it requests a major disaster declaration from the
president. Before that request is granted or denied, FEMA and its partners
conduct a *preliminary damage assessment* (PDA) that estimates the scale of
the damage and compares it against the statutory thresholds FEMA uses to gauge
whether federal assistance is warranted.

`get_preliminary_damage_assessments()` extracts these estimates from the PDF
PDA reports FEMA publishes and joins them to FEMA's own record of which
requests were granted and which were turned down, so that each report carries
its official outcome. We then ask a simple question: **do the requests FEMA
approves actually clear the Public Assistance per-capita threshold, and do the
ones it denies fall short?**

This is a step-through version of the package vignette
(`vignettes/preliminary_damage_assessments.Rmd.orig`). It runs the same
analysis, but it exposes the two switches that decide how long the first chunk
takes, and it adds sections for inspecting what the join produced. Run it a
chunk at a time; nothing below depends on being knitted as a whole.

## PDA Data Preparation

Because FEMA publishes only PDF reports of preliminary damage assessments---rather
than formally structuring these data, maintaining an ongoing database, and publishing
the data via an API---the package provides two functions for accessing these data: 

 - `scrape_pda_pdfs()` finds the URLs to every PDA report at 
  https://www.fema.gov/disaster/how-declared/preliminary-damage-assessments/reports and
  then systematically downloads these PDFs into a user-specified directory. These PDFs
  are the basis for the dataset produced by the following function. Because FEMA receives 
  and the President makes disaster request decisions on an ongoing basis, the results of this
  function quickly become stale: new PDA reports are available online but are not reflected
  in the user's specified directory. Accordingly, it is important to re-run this function 
  frequently if the most up-to-date data are important. The function does not re-download reports
  that are already saved in the user-specified directory, so after the initial run, which can be 
  slow, subsequent runs should be very fast--they only add new reports.
- `get_preliminary_damage_assessments()` walks the user's specified directory and extracts 
  structured attributes from each PDA report, compiling data from all reports in the directory 
  into a consistent dataset. 

## Data sources

| Role | Dataset | Access |
|---|---|---|
| Damage estimates | PDA reports | `climateapi::get_preliminary_damage_assessments()` |
| Approved requests | `DisasterDeclarationsSummaries` | OpenFEMA, via `rfema::open_fema()` |
| Denied requests | `DeclarationDenials` | OpenFEMA, via `rfema::open_fema()` |

The two FEMA datasets are *authoritative*: they are FEMA's own structured
records of what was declared and what was turned down. The PDAs, by contrast,
are text-extracted from unstructured PDFs and carry the caveats described in
`?get_preliminary_damage_assessments`. `get_preliminary_damage_assessments()`
treats the two authoritative datasets as defining the universe of requests and
maps the PDA reports back onto them, so that we can better understand the
relationship between preliminary damage estimates and disaster declaration
decisionmaking. Both FEMA datasets are fetched by the function itself; nothing
below calls the API directly.

```{r}
#| label: setup

library(climateapi)
library(tidyverse)
library(urbnthemes)

## `library(climateapi)` attaches the installed build. If that build predates the
## `join_outcomes` argument used below, either reinstall the package from this
## clone (devtools::install()) or, if the Imports are all present in your
## library, load the source directly instead:
# devtools::load_all()

set_urbn_defaults(style = "print")
```

### Two switches worth setting before you run the next chunk

`refresh_reports` controls whether FEMA's report listing is walked and any newly
published PDFs downloaded. It only adds reports that are not already on disk, so
after the first run it is quick, but it does need a network connection and
several minutes.

`rebuild_dataset` controls whether the PDFs are re-parsed. Parsing the full
archive takes a long time -- there are over 1,300 reports -- so leave it `FALSE`
while you are working through the analysis and set it `TRUE` when you have
downloaded new reports or changed the parsing code. The join to FEMA's
declaration and denial records is redone either way, because those records
change as FEMA publishes new decisions.

```{r}
#| label: switches

refresh_reports = FALSE
rebuild_dataset = FALSE
```

```{r}
#| label: load-pdas

if (refresh_reports) { scrape_pda_pdfs() }

pda_outcomes = get_preliminary_damage_assessments(use_cache = !rebuild_dataset)

dim(pda_outcomes)
```

## What the function does for us

Everything the join requires is done inside
`get_preliminary_damage_assessments()`, so this document does not repeat it. It
is worth stating what that work is, because the returned columns only make
sense alongside it.

**The rows are FEMA's, not the reports'.** The function returns one row for
every granted major disaster declaration and every turned-down major disaster
request FEMA publishes, going back to the start of both records. A request with
no usable PDA report keeps its row, with the report columns empty and
`pda_matched` set to `FALSE`.

**Approvals are matched on the disaster number.** It is unique nationally, so
it alone identifies the declaration. Where a declaration has more than one
report -- an original and an appeal -- the most recent is kept, and
`match_quality` says so.

**Denials are matched on the decision date, checked against the hazard.** A
denied request never receives a disaster number, so there is no key to join on.
What a denial report does carry is the date the decision was made, which is the
turndown's `request_status_date` in FEMA's record. That equality, within a
state, is the primary match. Where one date matches several of a state's
turndowns -- FEMA can decide two requests on the same day -- the candidates are
narrowed by how many hazard categories they share with the report, then by
elimination, since a turndown can back only one report. Anything still
ambiguous is left unmatched, counted in the function's consolidated warning,
rather than guessed at. A report
with no same-day turndown gets one further chance: the nearest earlier turndown
within seven days, and only when the hazards agree; those carry an
`"approximate"` `match_quality` so they stay visible.

The hazard comparison rests on two columns the function derives by keyword:
`hazards`, read from the report's own event title, and `denial_hazards`, read
from FEMA's incident name and requested incident types. Both use the same set
of category names ("flooding", "severe storm", "tornado", and so on), which is
what makes the two free-text descriptions comparable at all.

The identifying columns, so you can see the shape of a row before going
further:

```{r}
#| label: peek

pda_outcomes %>%
  select(
    disaster_number, state_name, decision, declaration_date,
    declaration_request_date, declaration_title, event_title, hazards,
    pda_matched, match_quality) %>%
  glimpse()
```

### Did the join succeed?

Before analyzing the joined data, we confirm that the PDAs actually matched an
authoritative FEMA record. A low match rate would mean our outcome labels are
unreliable.

```{r}
#| label: match-rate

pda_outcomes %>%
  count(decision, pda_matched) %>%
  pivot_wider(
    names_from = pda_matched, values_from = n,
    names_prefix = "matched_", values_fill = 0)
```

Most unmatched rows are older requests FEMA published no PDA report for, which
is the expected shape of a full-universe dataset rather than a failure. It does
mean that any analysis of the report fields has to filter on `pda_matched`.

`match_quality` says how each match was made. Anything other than an exact
match is worth reading before you rely on it.

```{r}
#| label: match-quality

pda_outcomes %>%
  filter(pda_matched) %>%
  count(match_quality, sort = TRUE)
```

### Auditing the denied matches by eye

Denials are the matches to check, because they are made on a date and a hazard
rather than on a key. This is every denied match with the evidence for it side
by side: the report's own title and hazards against FEMA's.

```{r}
#| label: denied-match-evidence

pda_outcomes %>%
  filter(pda_matched, decision == "Denied") %>%
  transmute(
    state_name,
    event_title,
    hazards,
    declaration_title,
    denial_hazards,
    requested_incident_types,
    event_date_determined,
    declaration_request_date,
    match_quality = str_extract(match_quality, "^[a-z]+")) %>%
  arrange(desc(event_date_determined)) %>%
  print(n = 25)
```

## Analysis: does the per-capita threshold predict the outcome?

FEMA's Public Assistance program compares a jurisdiction's estimated per-capita
damage against a statutory per-capita *indicator* (a dollar threshold that is
updated annually). The function supplies the ratio of the two as
`pa_threshold_ratio`: a value above 1.0 means the estimated damage exceeded the
threshold.

If the threshold drove the decision, approved requests would sit above the line
and denials below it.

```{r}
#| label: ratio-boxplot
#| fig-alt: "Boxplot with jittered points of the ratio of estimated statewide per-capita impact to FEMA's Public Assistance per-capita indicator, with a dashed line at a ratio of one."

pda_outcomes %>%
  filter(!is.na(pa_threshold_ratio)) %>%
  ggplot(aes(x = decision, y = pa_threshold_ratio, color = decision)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40") +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.5) +
  labs(
    title = "Do approved requests clear the Public Assistance per-capita threshold?",
    subtitle = "Ratio of estimated statewide per-capita impact to FEMA's statutory indicator",
    x = NULL,
    y = "Per-capita impact ÷ threshold") +
  theme(legend.position = "none")
```

The relationship is imperfect, which is the point: the per-capita indicator is
one input among several (insurance coverage, localized impacts, prior-year
disaster burden), so some sub-threshold requests are approved and some
above-threshold requests are denied. Plotting the ratio against the total cost
estimate shows how the two dimensions interact.

```{r}
#| label: ratio-cost-scatter
#| fig-alt: "Scatterplot of the per-capita threshold ratio against total estimated damage, with a dashed vertical line at a ratio of one."

pda_outcomes %>%
  filter(!is.na(pa_threshold_ratio), cost_estimate_ia_pa_total > 0) %>%
  ggplot(aes(
    x = pa_threshold_ratio,
    y = cost_estimate_ia_pa_total,
    color = decision)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
  geom_point(alpha = 0.7, size = 2) +
  scale_y_continuous(
    labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) +
  labs(
    title = "Threshold ratio versus total estimated damage",
    subtitle = "Combined Individual and Public Assistance cost estimate",
    x = "Per-capita impact ÷ threshold",
    y = "Total estimated cost",
    color = "Decision")
```

## Caveats

- **State recovery is best-effort.** The affected state is extracted from the
  report's free text and is set to the first state name that appears; a report
  that mentions a neighboring state first could be mislabeled. Approvals are
  unaffected, because they are matched on the disaster number alone, but a
  denial report with no recoverable state -- a tribal request names none --
  cannot be matched and its damage estimates do not appear.
- **Denied requests are matched by date, not by key.** Because turndowns lack a
  disaster number, the denied match uses the exact turndown decision date within
  a state, validated by hazard-category agreement, with ambiguous or
  hazard-inconsistent candidates left unmatched. A turndown FEMA has not
  yet published in `DeclarationDenials` therefore appears as an unmatched PDA
  rather than as a wrong match. The hazard categories themselves come from
  keyword matching on free text, so an unusual hazard description could leave a
  category unassigned (which weakens, but does not break, the validation).
- **`disaster_number` reliability.** For FEMA's newest filename convention
  (`FY25…`, `FY26…`), the disaster number is recovered from the PDF body rather
  than the filename; see `?get_preliminary_damage_assessments` for details.
- **Per-capita fields are text-extracted.** The `pa_per_capita_impact_*`
  fields come from parsing PDFs and can be missing or malformed for individual
  reports; rows with missing values are excluded from the plots.

## See also

- `get_preliminary_damage_assessments()`: the PDA extraction itself
- `get_fema_disaster_declarations()`: FEMA disaster declarations
- `get_public_assistance()`: obligated Public Assistance funding
- `get_ihp_registrations()`: Individual and Households Program registrations
