---
title: "Preliminary Damage Assessments and Declaration Outcomes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preliminary Damage Assessments and Declaration Outcomes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  dpi = 300,
  eval = TRUE,
  echo = TRUE)

options(scipen = 999)
```

## Overview

When a state, territory, or tribal government believes an event exceeds its
capacity to respond, it requests a major disaster declaration from the
president. Before that request is granted or denied, FEMA and its partners
conduct a *preliminary damage assessment* (PDA) that estimates the scale of
the damage and compares it against the statutory thresholds FEMA uses to gauge
whether federal assistance is warranted.

`get_preliminary_damage_assessments()` extracts these estimates from the
PDA reports FEMA publishes, and then joins them to authoritative datasets from
FEMA describing disaster declaration request approvals and denials. 

These combined data are valuable because they provide important information about
the characteristics of disasters--at an early stage, hence "preliminary"--that is
used directly to assess whether these disasters merit federal assistance. Unique attributes
available only from PDAs include estimates of damages to public infrastructure and housing,
along with counts of the numbers of housing units impacted, by level of impact. 

## Data sources

| Role | Dataset | Access |
|---|---|---|
| Damage estimates | PDA reports | `climateapi::get_preliminary_damage_assessments()` |
| Approved requests | `DisasterDeclarationsSummaries` | OpenFEMA, via `rfema::open_fema()` |
| Denied requests | `DeclarationDenials` | OpenFEMA, via `rfema::open_fema()` |

The two FEMA datasets are *authoritative*: they are FEMA's own structured
records of what was declared and what was turned down. The PDAs, by contrast,
are text-extracted from unstructured PDFs and carry the caveats described in
`?get_preliminary_damage_assessments`. In this piece, we treat observations 
from the authoritative FEMA datasets as definining the universe of declarations,
and we map PDAs back to this universe so that we can better understand the 
relationships between preliminary damages and disaster declaration decisionmaking. 

```{r setup}
library(climateapi)
library(tidyverse)
library(urbnthemes)

set_urbn_defaults(style = "print")
```

## PDA Data Preparation

Because FEMA publishes only PDF reports of preliminary damage assessments---rather
than formally structuring these data, maintaining an ongoing database, and publishing
the data via an API---the package provides two functions for accessing these data: 

 - `scrape_pda_pdfs()` finds the URLs to every PDA report at 
  https://www.fema.gov/disaster/how-declared/preliminary-damage-assessments/reports and
  then systematically downloads these PDFs into a user-specified directory. These PDFs
  are the basis for the dataset produced by the following function. Because FEMA receives 
  and the President makes disaster request decisions on an ongoing basis, the results of this
  function quickly become stale: new PDA reports are available online but are not reflected
  in the user's specified directory. Accordingly, it is important to re-run this function 
  frequently if the most up-to-date data are important. The function does not re-download reports
  that are already saved in the user-specified directory, so after the initial run, which can be 
  slow, subsequent runs should be very fast--they only add new reports.
- `get_preliminary_damage_assessments()` walks the user's specified directory and extracts 
  structured attributes from each PDA report, compiling data from all reports in the directory 
  into a consistent dataset. 

```{r load-pdas, eval = FALSE}
scrape_pda_pdfs()
pdas = get_preliminary_damage_assessments(use_cache = FALSE)
```

```{r load-pdas-hidden, echo = FALSE}
## this updates the local cache of all PDF reports
scrape_pda_pdfs()
## this extracts attributes from all reports
## use_cache = FALSE ensures the function doesn't simply return
## an older dataset (based on the previous set of PDF reports)
pdas = get_preliminary_damage_assessments(use_cache = FALSE)
```


## Analysis

With our data in hand, it's trivial to summarize the universe of disaster declaration requests
over time:

```{r}
pdas %>%
  ggplot() +
  geom_bar(aes(x = fema_decision_year, fill = fema_decision)) +
  scale_x_continuous(breaks = seq(1950, 2030, by = 10))
```

But access to systematic information on damages and housing impacts are novel, and
speak to how FEMA and presidents have made decisions over time.

```{r}
pdas %>%
  filter(pda_matched == 1, pda_pa_requested == 1) %>%
  ggplot(aes(x = fema_decision, y = log(pda_pa_threshold_ratio), color = fema_decision)) +
  geom_boxplot(fill = NA) +
  labs(
    title = "Many denied requests surpass the damages threshold used by FEMA" %>% str_wrap(130),
    x = "", 
    y = "Logged ratio of PA per-capita damages relative to the per-capita indidcator") 

pdas %>%
  filter(pda_matched == 1, pda_pa_requested == 1, fema_decision_year > 2007) %>%
  mutate(
    president = case_when(
      fema_decision_year %in% c(2008:2012) ~ "Obama 1",
      fema_decision_year %in% c(2013:2016) ~ "Obama 2",
      fema_decision_year %in% c(2017:2020) ~ "Trump 1",
      fema_decision_year %in% c(2021:2024) ~ "Biden 1",
      fema_decision_year %in% c(2025:2028) ~ "Trump 2"),
    president = factor(president, levels = c("Obama 1", "Obama 2", "Trump 1", "Biden 1", "Trump 2"), ordered = TRUE)) %>%
  ggplot(aes(x = fema_decision, y = log(pda_pa_threshold_ratio), color = fema_decision)) +
  geom_jitter(alpha = .2) +
  geom_boxplot(outliers = FALSE, fill = NA) +
  facet_wrap(~ president, nrow = 1) +
  labs(
    title = "Many denied requests surpass the damages threshold used by FEMA" %>% str_wrap(130),
    x = "", 
    y = "Logged ratio of PA per-capita damages relative to the per-capita indidcator") 
```

FEMA's Public Assistance program compares a jurisdiction's estimated per-capita
damage against a statutory per-capita *indicator* (a dollar threshold that is
updated annually). We express each PDA as the ratio of the two: a value above
1.0 means the estimated damage exceeded the threshold.
